---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SUSNEO Energy Dashboard

> A comprehensive Shiny application for energy consumption analysis and visualization

<!-- badges: start -->
[![CI](https://github.com/TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes/workflows/CI/badge.svg)](https://github.com/TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes/actions)
[![codecov](https://codecov.io/github/TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes/graph/badge.svg?token=C76NX21FLR)](https://codecov.io/github/TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes)
[![R-CMD-check](https://github.com/TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes/workflows/R-CMD-check/badge.svg)](https://github.com/TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes/actions)
<!-- badges: end -->

## Live Demo

**Try the live application**: [https://tiagoadrianunes.shinyapps.io/susneo-shiny-app/](https://tiagoadrianunes.shinyapps.io/susneo-shiny-app/)

## Overview

SUSNEO Energy Dashboard is an interactive energy dashboard built with R Shiny that provides comprehensive analysis and visualization of energy consumption data.

The application features:

- **Interactive Charts**: Time series, facility comparisons, and energy type distributions
- **KPI Metrics**: Real-time calculation of consumption, emissions, and efficiency ratios
- **Advanced Filtering**: Date ranges, facilities, and energy types
- **Responsive Design**: Modern UI with bslib and Bootstrap
- **Data Management**: CSV upload functionality with sample data

## Installation

### Prerequisites

Make sure you have R (>= 4.0.0) installed on your system.

### Install from GitHub

```{r, eval = FALSE}
# Install from GitHub using devtools
if (!require(devtools)) install.packages("devtools")
devtools::install_github("TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes")

# Or using pak (recommended)
if (!require(pak)) install.packages("pak")
pak::pak("TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes")
```


## Quick Start

### Launch the Application

```{r, eval = FALSE}
# Load the package
library(susneoEnergyDashboard)

# Run the Shiny application
run_app()
```

### Using Sample Data

The application includes sample energy consumption data for demonstration:

```{r, eval = FALSE}
# Access sample data
data("sample_data", package = "susneoEnergyDashboard")
head(sample_data)
```

### Data Format

Your CSV files should include these columns:

- `id`: Unique identifier
- `site`: Facility/location name
- `date`: Date in DD-MM-YYYY format
- `type`: Energy type (Electricity, Gas, Water, etc.)
- `value`: Energy consumption value
- `carbon_emission_in_kgco2e`: Carbon emissions (optional)

## Features

### Dashboard Components

- **KPI Cards**: Display key metrics including total consumption, emissions, daily averages, efficiency ratios, peak usage, and facility counts
- **Time Series Chart**: Interactive line chart showing energy consumption trends over time
- **Facility Comparison**: Column chart comparing total energy usage across different facilities
- **Data Table**: Detailed summary table with filtering and sorting capabilities

### Data Management

- **CSV Upload**: Import your own energy consumption data
- **Sample Data**: Built-in demonstration dataset
- **Data Validation**: Automatic validation of required columns and data types
- **Date Processing**: Flexible date parsing supporting multiple formats

### Filtering & Interactivity

- **Date Range Selection**: Filter data by custom date periods
- **Facility Selection**: Multi-select filtering by facility/site
- **Energy Type Selection**: Filter by specific energy types
- **Real-time Updates**: All charts and metrics update dynamically

## Usage Examples

### Basic Usage
```{r, eval = FALSE}
# Start the application
susneoEnergyDashboard::run_app()

# The app will open in your default browser
# Use "Load Sample Data" to see the demo
```

### Working with Custom Data
```{r, eval = FALSE}
# Prepare your data with required columns
my_data <- data.frame(
  id = 1:100,
  site = rep(c("Building_A", "Building_B"), 50),
  date = seq(as.Date("2024-01-01"), by = "day", length.out = 100),
  type = rep(c("Electricity", "Gas"), 50),
  value = rnorm(100, 1000, 200),
  carbon_emission_in_kgco2e = rnorm(100, 50, 10)
)

# Save as CSV and upload through the app interface
write.csv(my_data, "my_energy_data.csv", row.names = FALSE)
```

## Development

### Building from Source
```{r, eval = FALSE}
# Clone the repository
git clone https://github.com/TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes.git
cd susneo-shiny-tiago-adria-nunes

# Install development dependencies
devtools::install_deps(dependencies = TRUE)

# Load and test
devtools::load_all()
devtools::test()

# Run the app in development mode
devtools::load_all()
run_app()
```

### Package Structure

```
susneo/
â”œâ”€â”€ R/                          # R source code
â”‚   â”œâ”€â”€ app_*.R                # App configuration
â”‚   â”œâ”€â”€ mod_*.R                # Shiny modules
â”‚   â”œâ”€â”€ fct_*.R                # Business logic functions
â”‚   â”œâ”€â”€ utils_*.R              # Utility functions
â”‚   â””â”€â”€ class_data_manager.R   # Data management class
â”œâ”€â”€ data/                      # Package data
â”œâ”€â”€ tests/                     # Unit tests
â”œâ”€â”€ inst/                      # Package assets
â””â”€â”€ .github/workflows/         # CI/CD configuration
```

## Contributing

1. Fork the repository
2. Create a feature branch (`git checkout -b feature/amazing-feature`)
3. Commit your changes (`git commit -m 'Add amazing feature'`)
4. Push to the branch (`git push origin feature/amazing-feature`)
5. Open a Pull Request

## License

This project is licensed under the MIT License.

## Version Info

**Version**: `r golem::pkg_version()`
**Compiled**: `r Sys.time()`

## Development Status
```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Check if we're in development environment
in_dev <- file.exists("DESCRIPTION") && dir.exists("R")

if (in_dev) {
  # Development environment - load package
  tryCatch({
    devtools::load_all(quiet = TRUE)
    cat("**Package**: Development version loaded âœ…\n")

    # Test results summary
    test_results <- try({
      suppressMessages(devtools::test(quiet = TRUE))
      cat("**Tests**: All tests passing âœ…\n")
    }, silent = TRUE)

    if (inherits(test_results, "try-error")) {
      cat("**Tests**: See CI badges for test status\n")
    }

    # Coverage info (if available)
    coverage_info <- try({
      suppressMessages(covr::package_coverage(quiet = FALSE))
      coverage_pct <- round(covr::percent_coverage(coverage_info), 1)
      cat("**Coverage**: ", coverage_pct, "%\n", sep = "")
    }, silent = TRUE)

    if (inherits(coverage_info, "try-error")) {
      cat("**Coverage**: See CI badges for coverage status\n")
    }

  }, error = function(e) {
    cat("**Package**: Development environment detected\n")
    cat("**Tests**: Run `devtools::test()` to check\n")
    cat("**Coverage**: Run `covr::package_coverage()` to check\n")
  })
} else {
  # Check if package is installed
  pkg_installed <- try({
    suppressPackageStartupMessages(library(susneo, quietly = TRUE))
    cat("**Package**: Installed version loaded âœ…\n")
  }, silent = TRUE)

  if (inherits(pkg_installed, "try-error")) {
    cat("**Package**: Not installed - install from GitHub\n")
  }

  cat("**Tests**: See CI badges for current test status\n")
  cat("**Coverage**: See CI badges for current coverage\n")
}

# CI Status (always show)
cat("**CI Status**: See badges above for current build status\n")
cat("**Issues**: Check [GitHub Issues](https://github.com/TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes/issues)\n")
```

### Package Coverage
```{r, error = TRUE}
covr::package_coverage()
```

### Recent Updates

- âœ… Comprehensive test suite with 240+ tests
- âœ… Enhanced module test coverage (dashboard, KPI cards, data upload)
- âœ… CI/CD pipeline with multi-platform testing
- âœ… Code coverage tracking
- âœ… Automated linting and code quality checks
- âœ… Documentation with pkgdown
- âœ… Sample data and example usage
- âœ… Resolved namespace conflicts and import issues

### Known Issues

- ðŸ“ Check [Issues page](https://github.com/TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes/issues) for current status

### Roadmap

- [ ] Enhanced data visualization options
- [ ] Export functionality for charts and reports
- [ ] Advanced filtering and aggregation features
- [ ] Performance optimizations for large datasets
