---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SUSNEO Energy Dashboard

> A comprehensive Shiny application for energy consumption analysis and visualization

<!-- badges: start -->
[![CI](https://github.com/TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes/workflows/CI/badge.svg)](https://github.com/TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes/actions)
[![codecov](https://codecov.io/github/TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes/graph/badge.svg?token=C76NX21FLR)](https://codecov.io/github/TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes)
[![R-CMD-check](https://github.com/TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes/workflows/R-CMD-check/badge.svg)](https://github.com/TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes/actions)
<!-- badges: end -->

## Live Demo

**Try the live application**: [https://tiagoadrianunes.shinyapps.io/susneo-shiny-app/](https://tiagoadrianunes.shinyapps.io/susneo-shiny-app/)

## Overview

SUSNEO Energy Dashboard is an interactive energy dashboard built with R Shiny that provides comprehensive analysis and visualization of energy consumption data.

The application features:

- **Interactive Charts**: Time series and facility comparison charts
- **Linear Model Analysis**: CO2 emissions vs energy consumption correlation with statistical interpretation
- **KPI Metrics**: Real-time calculation of consumption, emissions, and efficiency ratios
- **Advanced Filtering**: Date ranges, facilities, and energy types
- **Responsive Design**: Modern UI with bslib and Bootstrap
- **Data Management**: Loads the dataset shipped in the package's `data/` folder and cleans it automatically

## Installation

### Prerequisites

Make sure you have R (>= 4.1.0) installed on your system.

### Install from GitHub

```{r, eval = FALSE}
# Install from GitHub using devtools
if (!require(devtools)) install.packages("devtools")
devtools::install_github("TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes")

# Or using pak (recommended)
if (!require(pak)) install.packages("pak")
pak::pak("TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes")
```


## Quick Start

### Launch the Application

```{r, eval = FALSE}
# Load the package
library(susneoEnergyDashboard)

# Run the Shiny application
run_app()
```

### Using Sample Data

The application includes sample energy consumption data for demonstration:

```{r, eval = FALSE}
# Access sample data
data("sample_data", package = "susneoEnergyDashboard")
head(sample_data)
```

### Data Format

The dashboard reads the `sample_data` dataset from the package's `data/` folder (`data/sample_data.csv`). It has these columns:

- `id`: Unique identifier
- `site`: Facility/location name
- `date`: Date in DD-MM-YYYY format
- `type`: Energy type (Electricity, Gas, Water, etc.)
- `value`: Energy consumption value
- `carbon_emission_in_kgco2e`: Carbon emissions (optional)

Things to know about the data:

- **Dates are read day first.** A numeric date such as `03-04-2025` is read as 3 April (DD-MM-YYYY); layouts such as `2025-04-03` are accepted too.
- **Unusable rows are dropped** when the data is loaded: rows with an unreadable date or a non-numeric `value`. A missing `carbon_emission_in_kgco2e` counts as 0.
- **`value` has no unit.** Electricity, gas, water and the other types are measured in different units. When a selection mixes energy types the dashboard shows a warning, because the totals, charts and the linear model add those values together as they are. Select a single energy type for a like-for-like comparison.

## Features

### Dashboard Components

- **KPI Cards**: Display key metrics including total consumption, emissions, daily averages, efficiency ratios, peak usage, and facility counts
- **Time Series Chart**: Interactive line chart showing energy consumption trends over time
- **Facility Comparison**: Column chart comparing total energy usage across different facilities
- **Linear Model Analysis**: Statistical analysis of CO2 emissions vs energy consumption relationship with:
  - Interactive scatter plot with regression line
  - Model summary table (coefficients, R-squared, F-statistic)
  - Plain-language interpretation of statistical significance
  - Relationship strength and direction analysis
- **Data Table**: Detailed summary table with filtering and sorting capabilities
- **Mixed-units warning**: Shown above the KPI cards whenever the selection combines several energy types

### Data Management

- **Sample Data**: Built-in dataset loaded from the package's `data/` folder
- **Data Cleaning**: Dates are parsed and values checked; unusable rows are dropped
- **Date Processing**: Flexible date parsing supporting multiple formats; ambiguous numeric dates are read day first (DD-MM-YYYY)

### Filtering & Interactivity

- **Date Range Selection**: Filter data by custom date periods
- **Facility Selection**: Multi-select filtering by facility/site
- **Energy Type Selection**: Filter by specific energy types
- **Real-time Updates**: All charts and metrics update dynamically

## Usage Examples

### Basic Usage
```{r, eval = FALSE}
# Start the application
susneoEnergyDashboard::run_app()

# The app opens in your default browser and shows the sample data
```

## Development

### Building from Source
```{r, eval = FALSE}
# Clone the repository
git clone https://github.com/TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes.git
cd susneo-shiny-tiago-adria-nunes

# Install development dependencies
devtools::install_deps(dependencies = TRUE)

# Load and test
devtools::load_all()
devtools::test()

# Run the app in development mode
devtools::load_all()
run_app()
```

### Package Structure

```
susneo/
├── R/                          # R source code
│   ├── app_*.R                # App configuration
│   ├── mod_*.R                # Shiny modules
│   ├── class_*.R              # R6 class that owns the app's data (EnergyDataManager)
│   ├── fct_*.R                # Business logic as pure functions (data, charts, KPIs, linear model)
│   └── utils_*.R              # Utility functions
├── data/                      # Package data
├── tests/                     # Unit tests
├── inst/                      # Package assets
└── .github/workflows/         # CI/CD configuration
```

## Contributing

1. Fork the repository
2. Create a feature branch (`git checkout -b feature/amazing-feature`)
3. Commit your changes (`git commit -m 'Add amazing feature'`)
4. Push to the branch (`git push origin feature/amazing-feature`)
5. Open a Pull Request

## License

This project is licensed under the MIT License.

## Version Info

**Version**: `r golem::pkg_version()`
**Compiled**: `r Sys.time()`

## Development Status
```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Check if we're in development environment
in_dev <- file.exists("DESCRIPTION") && dir.exists("R")

if (in_dev) {
  # Development environment - load package
  tryCatch({
    devtools::load_all(quiet = TRUE)
    cat("Package: Development version loaded ✅\n")

    # Test results summary
    test_results <- try({
      suppressMessages(devtools::test(quiet = TRUE))
      cat("Tests: All tests passing ✅\n")
    }, silent = TRUE)

    if (inherits(test_results, "try-error")) {
      cat("Tests: See CI badges for test status\n")
    }

    # Coverage info (if available)
    coverage_info <- try({
      suppressMessages(covr::package_coverage(quiet = FALSE))
      coverage_pct <- round(covr::percent_coverage(coverage_info), 1)
      cat("Coverage: ", coverage_pct, "%\n", sep = "")
    }, silent = TRUE)

    if (inherits(coverage_info, "try-error")) {
      cat("Coverage: See CI badges for coverage status\n")
    }

  }, error = function(e) {
    cat("Package: Development environment detected\n")
    cat("Tests: Run `devtools::test()` to check\n")
    cat("Coverage: Run `covr::package_coverage()` to check\n")
  })
} else {
  # Check if package is installed
  pkg_installed <- try({
    suppressPackageStartupMessages(library(susneo, quietly = TRUE))
    cat("Package: Installed version loaded ✅\n")
  }, silent = TRUE)

  if (inherits(pkg_installed, "try-error")) {
    cat("Package: Not installed - install from GitHub\n")
  }

  cat("Tests: See CI badges for current test status\n")
  cat("Coverage: See CI badges for current coverage\n")
}

# CI Status (always show)
cat("CI Status: See badges above for current build status\n")
```

### Package Coverage
```{r, error = TRUE}
covr::package_coverage()
```

### Recent Updates
#### Version `r golem::pkg_version()`
- **Fixed**: the linear model no longer fails when emissions do not vary; it needs at least 3 complete rows, and the p-value explanation is corrected
- **Fixed**: large round totals are no longer shown in scientific notation (for example `1e+06`)
- **Added**: warning when a selection mixes energy types (they can use different units)
- **Changed**: the R6 class is now `EnergyDataManager`, which owns the cleaned data and serves the dashboard's filters. It delegates the calculations to plain functions (`fct_data.R`, `fct_linear_model.R`) that can be tested without Shiny. All package code uses `importFrom` instead of `pkg::fn`
- **Removed**: the CSV upload and "Load Sample Data" code, which was never reachable from the app, and the energy type pie chart and trend chart, which were never displayed

#### Version 0.0.0.9010
- **Linear Model Analysis**: New statistical analysis module for CO2 emissions vs energy consumption
  - Interactive scatter plot with regression line visualization
  - Comprehensive model summary table with statistical metrics
  - Plain-language interpretation of results and significance levels
  - Modular architecture following best practices
  - Full test coverage with realistic data scenarios

#### Previous Updates
- Comprehensive test suite with 320+ tests
- Enhanced module test coverage (dashboard, KPI cards, data upload, linear model)
- CI/CD pipeline with multi-platform testing
- Code coverage tracking
- Automated linting and code quality checks
- Documentation with pkgdown
- Sample data and example usage
- Resolved namespace conflicts and import issues
- Adjusted the Value Boxes

### Known Issues

- Check [Issues page](https://github.com/TiagoAdriaNunes/susneo-shiny-tiago-adria-nunes/issues) for current status

### Roadmap
- [x] Linear model analysis for CO2 vs energy correlation ✅ (v0.0.0.9010)
- [ ] Enhanced data visualization options
- [ ] Export functionality for charts and reports
- [ ] Advanced statistical models (multiple regression, time series analysis)
- [ ] Advanced filtering and aggregation features
- [ ] Performance optimizations for large datasets
